---
title: "Assignment 07: Web Scraping the FSB Directory"
author: "Marina Erceg"
date: "2026-10-06"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true 
    code_download: true 
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
```

## Goal

Build one data frame with every faculty/staff member from all seven FSB departments, with four columns: department, name, position, and website. Print it with `head()`.

## Step 1: Check that scraping is allowed

```{r checking_webpage_scrapability}
robotstxt::paths_allowed(
  paths = "/fsb/directory/?up=/query/all/all/Information%20Systems%20--%20Analytics/all",
  domain = "miamioh.edu"
)
```

## Step 2: Get one department working first

```{r loading_isa_directory_page}
url_isa = "https://miamioh.edu/fsb/directory/?up=/query/all/all/Information%20Systems%20--%20Analytics/all"

page_isa = rvest::read_html(url_isa)

all_rows_isa = page_isa |> rvest::html_elements("tr")
people_isa = all_rows_isa[-1]

length(people_isa)
```

```{r building_isa_table}
isa_department = people_isa |> rvest::html_element("td:nth-child(1)") |> rvest::html_text2()
isa_name = people_isa |> rvest::html_element("strong a") |> rvest::html_text2()
isa_position = people_isa |> rvest::html_element("i") |> rvest::html_text2() |> stringr::str_squish()
isa_website = people_isa |> rvest::html_element("strong a") |> rvest::html_attr("href")

isa_df = tibble::tibble(
  department = isa_department,
  name = isa_name,
  position = isa_position,
  website = isa_website
)
```

```{r previewing_isa_table}
head(isa_df)

colSums(is.na(isa_df))
```

## Step 3: Turn it into a function

```{r defining_scrape_fsb_dept_function}
scrape_fsb_dept = function(dept_query) {

  dept_url = paste0(
    "https://miamioh.edu/fsb/directory/?up=/query/all/all/",
    utils::URLencode(dept_query),
    "/all"
  )

  Sys.sleep(1)

  dept_page = rvest::read_html(dept_url)
  dept_all_rows = dept_page |> rvest::html_elements("tr")
  dept_people = dept_all_rows[-1]

  dept_df = tibble::tibble(
    department = dept_people |> rvest::html_element("td:nth-child(1)") |> rvest::html_text2(),
    name = dept_people |> rvest::html_element("strong a") |> rvest::html_text2(),
    position = dept_people |> rvest::html_element("i") |> rvest::html_text2() |> stringr::str_squish(),
    website = dept_people |> rvest::html_element("strong a") |> rvest::html_attr("href")
  )

  return(dept_df)
}
```

```{r testing_function_on_isa}
test_isa = scrape_fsb_dept("Information Systems -- Analytics")

head(test_isa)
```

## Step 4: Run it on all seven departments

```{r scraping_all_seven_departments}
fsb_dept_queries = c(
  "Accountancy",
  "Economics",
  "Entrepreneurship",
  "Finance",
  "Information Systems -- Analytics",
  "Marketing",
  "Management"
)

fsb_directory_raw = purrr::map_df(.x = fsb_dept_queries, .f = scrape_fsb_dept)
```

The directory lists some people under a different department than the one searched (for example, a Marketing search returns people from the First-Year Integrated Core), and some people show up in more than one search. This keeps only the seven departments and drops exact duplicate rows.

```{r keeping_seven_departments_and_removing_duplicates}
fsb_seven_departments = c(
  "Accountancy",
  "Economics",
  "Entrepreneurship",
  "Finance",
  "Information Systems & Analytics",
  "Marketing",
  "Management"
)

fsb_directory = fsb_directory_raw |>
  dplyr::filter(department %in% fsb_seven_departments) |>
  dplyr::distinct()
```

## Step 5: Check the result before knitting

```{r checking_final_data_quality}
dplyr::count(fsb_directory, department)

colSums(is.na(fsb_directory))

nrow(fsb_directory_raw)
nrow(fsb_directory)

fsb_directory |> dplyr::count(name) |> dplyr::filter(n > 1)
```

## Final output

```{r printing_final_head}
head(fsb_directory) |> print(width = Inf)
```
