---
title: "airbnb_analysis"
author: "Marina Erceg"
date: "2026-09-28"
output: 
  html_document:
    toc: true
    toc_float: true
    numbered_sections: true
    code_folding: show
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read_in_data}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/midwest_airbnb_extended.db")
listings     = DBI::dbReadTable(con, "listings") |> tibble::as_tibble()
availability = DBI::dbReadTable(con, "availability_monthly") |> tibble::as_tibble()
reviews      = DBI::dbReadTable(con, "reviews_monthly") |> tibble::as_tibble()
DBI::dbDisconnect(con)
```

## Task 1
Who owns the market: listings per host, joined back to the listings.
```{r task_one}
host_counts = listings |>
  dplyr::count(host_id, name = "host_listings")

host_counts |>
  dplyr::filter(host_listings >= 10) |>
  dplyr::summarize(hosts_10_plus = dplyr::n())

listings_hosts = listings |>
  dplyr::left_join(host_counts, by = "host_id")

listings_hosts |>
  dplyr::group_by(city) |>
  dplyr::summarize(
    listings = dplyr::n(),
    pct_held_by_10_plus_hosts = 100 * mean(host_listings >= 10)
  ) |>
  as.data.frame()
```

## Task 2
Value for a family of four.
```{r task_two}
family = listings |>
  dplyr::filter(
    room_type == "Entire home/apt",
    accommodates >= 4,
    number_of_reviews >= 10,
    review_scores_rating >= 4.8
  ) |>
  dplyr::mutate(price_per_person = price / accommodates) |>
  dplyr::group_by(city, neighbourhood) |>
  dplyr::summarize(
    n = dplyr::n(),
    median_price_per_person = median(price_per_person, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::filter(n >= 20)

family |>
  dplyr::filter(city == "Columbus") |>
  dplyr::arrange(median_price_per_person)
```

## Task 3
What earns: medians by city and room type in one summarize call.
```{r task_three}
earnings = listings |>
  dplyr::group_by(city, room_type) |>
  dplyr::summarize(
    dplyr::across(
      c(price, reviews_per_month, estimated_revenue_l365d),
      ~ median(.x, na.rm = TRUE)
    ),
    .groups = "drop"
  )

earnings |> print(n = Inf)
```

## Task 4
Concentration: share of each city’s revenue earned by its top 10 percent of listings.
```{r task_four}
concentration = listings |>
  dplyr::group_by(city) |>
  dplyr::arrange(dplyr::desc(estimated_revenue_l365d), .by_group = TRUE) |>
  dplyr::mutate(
    rank = dplyr::row_number(),
    top10 = rank <= ceiling(0.10 * dplyr::n())
  ) |>
  dplyr::summarize(
    pct_revenue_top10 = 100 * sum(estimated_revenue_l365d[top10], na.rm = TRUE) /
      sum(estimated_revenue_l365d, na.rm = TRUE)
  )

concentration
```

## Task 5
The year ahead: October 2026 availability for entire homes.
```{r task_five}
entire = listings |>
  dplyr::filter(room_type == "Entire home/apt")

october = availability |>
  dplyr::filter(month == "2026-10")

entire |>
  dplyr::inner_join(october, by = "id") |>
  dplyr::mutate(mostly_taken = nights_available < 0.5 * nights_in_month) |>
  dplyr::group_by(city) |>
  dplyr::summarize(
    homes = dplyr::n(),
    pct_mostly_taken = 100 * mean(mostly_taken, na.rm = TRUE)
  ) |>
  dplyr::arrange(pct_mostly_taken)
```

## Task 6
Demand by season: total reviews per city and month, July 2025 to June 2026.
```{r task_six}
seasonal = reviews |>
  dplyr::filter(month >= "2025-07", month <= "2026-06") |>
  dplyr::inner_join(dplyr::select(listings, id, city), by = "id") |>
  dplyr::group_by(city, month) |>
  dplyr::summarize(total_reviews = sum(reviews), .groups = "drop")

seasonal |>
  dplyr::group_by(city) |>
  dplyr::summarize(
    busiest_month = month[which.max(total_reviews)],
    quietest_month = month[which.min(total_reviews)],
    busiest_to_quietest = max(total_reviews) / min(total_reviews)
  )
```

## Task 7 
The Superhost premiums for entire homes.
```{r task_seven}
superhost = entire |>
  dplyr::filter(host_is_superhost %in% c("t", "f")) |>
  dplyr::group_by(city, host_is_superhost) |>
  dplyr::summarize(
    listings = dplyr::n(),
    median_price = median(price, na.rm = TRUE),
    median_rating = median(review_scores_rating, na.rm = TRUE),
    .groups = "drop"
  )

superhost

superhost |>
  dplyr::group_by(city) |>
  dplyr::summarize(
    premium_dollars = median_price[host_is_superhost == "t"] -
      median_price[host_is_superhost == "f"]
  )
```

## Task 8
What the join lost: listings with no rows in availability.
```{r task_eight}
no_calendar = listings |>
  dplyr::anti_join(availability, by = "id")

no_calendar |> dplyr::summarize(listings_without_availability = dplyr::n())

no_calendar |> dplyr::count(city, sort = TRUE)
```